---
title: "BrainMaps"
author: "Valerie Jill Sydnor"
output: 
  html_document:
    theme: flatly
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(reshape2)
library(matrixStats)
library(ggseg)
library(ggsegExtra)
library(ggsegGlasser)
```


```{r}
brains <- read.csv("./brainmaps_glasser.csv", header=TRUE) #read in cortical map data
regions <- read.csv("./glasser_regions.csv", header=TRUE) #read in glasser parcel names
```

**Derive Axis Rankings**

```{r}
brains.ranked <- lapply(brains, function(x){ rank(x) }) #for each of the 10 cortical measures, derive unidimensional axis orderings by ranking by parcel value
brains.ranked <- as.data.frame(brains.ranked) #format as df
brains.ranked$T1T2ratio <- rank(-brains.ranked$T1T2ratio) # reverse order T1-weighted to T2-weighted ratio rankings so that sensorimotor regions are ranked lower (higher T1w/T2w ratio) and association regions are ranked higher (lower T1w/T2w ratio)
brains.ranked$Externopyramidisation <- rank(-brains.ranked$Externopyramidisation) # reverse order Externopyramidisation rankings so that sensorimotor regions are ranked lower (higher externopyramidisation) and association regions are ranked higher (higher internopyramidisation)
brains.ranked <- cbind(regions,brains.ranked) #add parcel names to df
```

**Visualize Cortical Maps**

*Anatomical Hierarchy*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=T1T2ratio)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/T1T2Ratio.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```

*Functional Hierarchy*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=G1.fMRI)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/G1fMRI.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```
*Evolutionary Hierarchy*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=Evolution.Expansion)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/EvolutionExpansion.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```

*Allometric Scaling*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=AllometricScaling.PNC20mm)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/AllometricScaling.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```

*Aerobic Glycolysis*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=PET.AG)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/PETAG.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```

*Cerebral Blood Flow*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=CBF)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/CBF.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```

*Gene Expression*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=PC1.AHBA)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/PC1AHBA.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
``` 

*NeuroSynth*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=PC1.Neurosynth)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/PC1Neurosynth.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```

*Externopyramidization*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=Externopyramidisation)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/Externopyramidisation.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```

*Cortical Thickness*
```{r, message = F, warning = F}
ggseg(.data = brains.ranked, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=Cortical.Thickness)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", name = NULL, midpoint = 90)

ggsave(filename = "./visualizations/CorticalThickness.pdf", device = "pdf", dpi = 300, width = 4 , height = 4)
```


**Visualize Archetypal Sensorimotor-Association Axis**

```{r, message = F, warning = F}
brains.average.rank <- brains.ranked %>% dplyr::select(-region)
brains.average.rank <- rowMeans(brains.average.rank) #average all 10 cortical rankings to get each parcel's average axis rank
brains.average.rank <- cbind(regions, brains.average.rank) #add glasser parcel names to average axis rankings and convert to df
brains.average.rank$final.rank <- rank(brains.average.rank$brains.average.rank) #rerank parcels from 1 to 180 by average axis rank to visualize the archetypal sensorimotor-association axis 
write.csv(brains.average.rank, file="./Sensorimotor_Association_Axis_AverageRanks.csv", row.names = FALSE) #save out average ranks

ggseg(.data = brains.average.rank, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=final.rank)) + theme_void() + scale_fill_gradient2(low= "goldenrod1", mid = "white", high = "#6f1282", guide = "colourbar", aesthetics = "fill", midpoint = 90)

ggsave(filename = "./visualizations/AverageSAAxis.pdf", device = "pdf", dpi = 300, width = 6 , height = 6)
```

**Visualize Rank Variability**

```{r, message = F, warning = F}
brains.MAD.rank <- brains.ranked %>% dplyr::select(-region)
brains.MAD.rank <- rowMads((as.matrix(brains.MAD.rank))) #calculate parcel median absolute deviation
brains.MAD.rank <- cbind(regions, brains.MAD.rank) #add glasser parcel names to MAD and convert to df

ggseg(.data = brains.MAD.rank, atlas = "glasser", hemisphere = c("left"), mapping=aes(fill=brains.MAD.rank)) + theme_void() +  scale_fill_gradientn(colors = c("#2B1C4F","#283063","#27376A","#264578","#26568B","#2B5F99","#467ebd","#6399d6","#a1caf7","#bcd8f7","#d4e8ff"))

ggsave(filename = "./visualizations/SAAxisVariability.pdf", device = "pdf", dpi = 300, width = 6 , height = 6)
```

**Sensorimotor and Association Tertile Average Feature Z-scores**

```{r}
brains.tertiles <- cbind(brains, brains.average.rank)
brains.tertiles <-  brains.tertiles[
order(brains.tertiles[,13] , decreasing = FALSE), #order data for all 10 cortical measures by parcel final rank
] 
brains.tertiles <- brains.tertiles %>% dplyr::select(-region, -brains.average.rank, -final.rank)
brains.tertiles <- lapply(brains.tertiles, function(x){ scale(x) }) #z-score data for each of the 10 cortical measures
brains.tertiles <- as.data.frame(brains.tertiles) #format as df

brains.lowertertile <- brains.tertiles[1:60,] #lower tertile (sensorimotor tertile)
brains.uppertertile <- brains.tertiles[121:180,] #upper tertile (association tertile)

lowertertile.mean <- lapply(brains.lowertertile, function(x){ mean(x) }) #calculate sensorimotor tertile average z-score for each of the 10 cortical measures
uppertertile.mean <- lapply(brains.uppertertile, function(x){ mean(x) }) #calculate association tertile average z-score for each of the 10 cortical measures

tertile.average.zscores <- as.data.frame(rbind(lowertertile.mean, uppertertile.mean)) 
tertile.average.zscores[1,2] <- uppertertile.mean$T1T2ratio #switch lower and upper T1-weighted/T2-weighted ratio z-scores to correctly reflect sensorimotor versus association tertile
tertile.average.zscores[2,2] <- lowertertile.mean$T1T2ratio #switch lower and upper T1-weighted/T2-weighted ratio z-scores to correctly reflect sensorimotor versus association tertile
tertile.average.zscores[1,7] <- uppertertile.mean$Externopyramidisation #switch lower and upper Externopyramidisation z-scores to correctly reflect sensorimotor versus association tertile
tertile.average.zscores[2,7] <- lowertertile.mean$Externopyramidisation #switch lower and upper Externopyramidisation z-scores to correctly reflect sensorimotor versus association tertile

tertile.zscores.plotdata <- as.data.frame((t(tertile.average.zscores))) #transpose df
tertile.zscores.plotdata <- tertile.zscores.plotdata[c(2, 1, 3, 5, 6, 4, 9, 10, 7, 8),] #reorder rows as AH, FH, EH, AS, AG, CB, GE, NS, EX, CT
tertile.zscores.plotdata$brainmap <- rownames(tertile.zscores.plotdata) #add brainmap identifier column 
tertile.zscores.plotdata$lowertertile.mean <- as.numeric(tertile.zscores.plotdata$lowertertile.mean)
tertile.zscores.plotdata$uppertertile.mean <- as.numeric(tertile.zscores.plotdata$uppertertile.mean)

tertile.zscores.plotdata <- melt(tertile.zscores.plotdata, id.vars=c("brainmap"), measure.vars=c("lowertertile.mean","uppertertile.mean"), variable.name = "tertile", value.name="zscore") #convert to long format for plotting

ggplot(tertile.zscores.plotdata, aes(x=brainmap, y=zscore, fill=tertile)) + 
  geom_point(color="white", shape=22, size=13) + 
  labs(x="\n", y="\n") +
  scale_x_discrete(limits=tertile.zscores.plotdata$brainmap) +
  theme(panel.background = element_blank(), panel.grid = element_line(color = "gray92"),  axis.text.x = element_text(angle = 90), legend.position = "none") +
  coord_cartesian(xlim=c(1,10)) +
  scale_y_continuous(limits=c(-1.2,1.2)) +
  scale_fill_manual(values=c("goldenrod1", "#6f1282"), labels=c("sensorimotor tertile","association tertile")) 

ggsave(filename = "./visualizations/SAAxis_Tertile_Zscores.pdf", device = "pdf", dpi = 300, width = 6 , height = 6)
```
